Reanalysis Step 7: Sample Comparison Table

### 2025-09-04 reanalysis-7-missingness_patterns.qmd
data comparison

# 2. Load new complete cases data (from reanalysis-1)
complete_file <- file.path(reanalysis_data_dir, "df_complete_original.rds")
if (!file.exists(complete_file)) {
  stop("Complete cases data not found. Run reanalysis-1-data_preparation.qmd first.")
}
complete_data <- readRDS(complete_file)

# Split into submitted and excluded based on cohort_id
submitted_data <- complete_data[complete_data$cohort_id == "submitted", ]
excluded_data <- complete_data[complete_data$cohort_id == "excluded", ]

@@

# Variables in original data (with proper capitalization)
vars_categorical_original <- c("gender_cat", "race_ethnic_cat", "insurance_cat", "age_cat",
                               "ever_lapse_binary", 
                               "Glaucoma_bef_hitplus_cat", "Otherretina_bef_hitplus_cat",
                               "catsurg_before_hitplus_cat",
                               "anti_VEGF", "PRP_flag", "focal_laser_flag", "other_inject",
                               "outcome_va_vi_binary")

@@

# All categorical variables for new data
vars_categorical_new <- c("gender_cat", "race_ethnic_cat", "insurance_cat", "age_cat",
                          "dr_severity", "ever_lapse_binary", 
                          "Glaucoma_bef_hitplus_cat", "Otherretina_bef_hitplus_cat",
                          "catsurg_before_hitplus_cat",
                          "anti_VEGF", "PRP_flag", "focal_laser_flag", "other_inject",
                          "outcome_va_vi_binary")

@@

  # Categorical variables
  cat_summaries <- do.call(rbind, lapply(categorical_vars, function(v) {
    if (v %in% names(data)) {
      summarize_categorical(data, v)
    } else {
      data.frame(variable = v, n = NA, pct = NA, n_missing = NA, pct_missing = NA)
    }
  }))

@@

>>>>

# 2. Load full data for imputation (contains both submitted and excluded)
imputation_data_file <- file.path(reanalysis_data_dir, "df_for_imputation.rds")
if (!file.exists(imputation_data_file)) {
  stop("Imputation data not found. Run reanalysis-1-data_preparation.qmd first.")
}
imputation_data <- readRDS(imputation_data_file)

# Split into submitted and excluded based on cohort_id
submitted_data <- imputation_data[imputation_data$cohort_id == "submitted", ]
excluded_data <- imputation_data[imputation_data$cohort_id == "excluded", ]

@@

# Variables in original data (with proper capitalization)
vars_categorical_original <- c("gender_cat", "race_ethnic_cat", "insurance_cat", "age_cat",
                               "ever_lapse_binary", 
                               "glaucoma_bef_hitplus_cat", "otherretina_bef_hitplus_cat",
                               "catsurg_before_hitplus_cat",
                               "anti_VEGF", "PRP_flag", "focal_laser_flag", "other_inject",
                               "outcome_va_vi_binary")

@@

# All categorical variables for new data
vars_categorical_new <- c("gender_cat", "race_ethnic_cat", "insurance_cat", "age_cat",
                          "dr_severity", "ever_lapse_binary", 
                          "glaucoma_bef_hitplus_cat", "otherretina_bef_hitplus_cat",
                          "catsurg_before_hitplus_cat",
                          "anti_VEGF", "PRP_flag", "focal_laser_flag", "other_inject",
                          "outcome_va_vi_binary")

@@

  # Categorical variables - filter out variables that don't exist
  cat_summaries_list <- lapply(categorical_vars, function(v) {
    if (v %in% names(data)) {
      summarize_categorical(data, v)
    } else {
      NULL
    }
  })

  # Remove NULL entries and combine
  cat_summaries_list <- cat_summaries_list[!sapply(cat_summaries_list, is.null)]

  if (length(cat_summaries_list) > 0) {
    cat_summaries <- do.call(rbind, cat_summaries_list)
  } else {
    cat_summaries <- data.frame(variable = character(), n = numeric(), 
                                pct = numeric(), n_missing = numeric(), 
                                pct_missing = numeric())
  }

@@

**Note:** The "Excluded" column shows patients who were excluded from the original submitted analysis because they lacked 2-year outcome data. These patients have substantial missing data, which is why many cells show 0 (NaN%) - indicating all values are missing for those variables. The imputation process fills in these missing values, allowing us to include these patients in the analysis.
